---
title: "The datafinder pipeline"
author: "Sean Raleigh and Jack Wilburn"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{The datafinder pipelin}e
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Thanks for choosing to use datafinder as an exploratory tool for your next project! As with any package, there are many functions and the best way to acquaint yourself with them is through this documentation. Here we will discuss the entire datafinder pipeline from project definition to model building.

Let's make sure that the package is loaded and attached with:

```{r, warning = FALSE, message = FALSE}
library(datafinder)
```

Also, we expect that you have a project in mind. For this example, we are going to be looking for data that will give us a 2 sample t-test that has a significant p value. This might be useful for statistical educators but if your project is more research based then make sure that the data you're using satisfies all necessary conditions.

## Discovering Loaded Pacakges

The first thing we need to do is see which packages we have installed have data in them. We do this by running the `discover_packages()` command from the terminal. 

```{r, warning = FALSE, message = FALSE}
tail(discover_packages(), 10)
```

Running this command on my machine yields a large list of packages with the amount of data sets in each. The command above grabs just the last 10 of them. For this example I'm going to explore the datasets package but for your project you may choose another. Additionally, you may have to explore a few before you find the right data.

## Exploring A Single Package

Now let's explore the datasets package. It's really easy to look at the data in the package with the `explore_package()` function.

```{r}
explore_package(datasets)[67:73,]
```

Again, this function gives a large list of datasets so we're just showing some of them. For this project we're going to choose the mtcars data set.

## Visualizing The Data Frame

Now it's time to look at the whole dataframe and get a sense of the data contained within. to do this, we can use the `visualize_dataframe()` command. This command can sometimes take a while to run but rest assured it's running.

```{r, fig.width = 5, fig.height = 5}
visualize_dataframe(mtcars)
```

This plot is giving us a 10 by 10 grid which is the max output for this function. If you have some more variables it might be best to use the overide to plot more variables. The syntax for the override would be `visualize_dataset(mtcars, override = TRUE)`.



